Parallel Information Set Generation for Kriegspiel

### 2012-06-27 sbac/paper.tex
Add Sanjay as author
remoived dumplicatecoinclusion.. some reorganiuzation

\author{Mark Richards \and Abhishek Gupta \and Osman Sarood \\
%Department of Computer Science\\ 
%University of Illinois\\ 
%Urbana, IL\\ mdrichar@illinois.edu\\
% For a paper whose authors are all at the same institution, 
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'', 
% just like the second author.
%\and
%Second Author\\
%Institution2\\
%First line of institution2 address\\ Second line of institution2 address\\ 
%SecondAuthor@institution2.com\\
}

@@

\Section{Background}\label{Background}
\subsection{Kriegspiel}
Kriegspiel, or partially observable chess,  was invented by Henry Michael Temple in 1899.  Historically, the
game has required three parties: two competing players and one referee.  Today,
players often compete over a computer network, with the role of the referee
being automated.  One player sees and controls only the white pieces; the
other, black.  The referee track of both sets of pieces.  The players alternate
moves as in a regular chess game, but instead of openly declaring their moves,
they secretly communicate move requests to the referee.  The referee, who knows
the location of all pieces, checks the legality of the move.

@@

\SubSection{Example}
\begin{figure}[ht]
\centering
\begin{minipage}[b]{0.5\linewidth}
\includegraphics[width=0.7\textwidth]{images/9B.png}
\caption{Actual Position}
\label{fig:figure1}
\end{minipage}
\hspace{0.5cm}
\centering
\begin{minipage}[b]{0.5\linewidth}
\includegraphics[width=0.7\textwidth]{images/OneView.png}
\caption{White's view}
\label{fig:figure2}
\end{minipage}
\end{figure}

@@

\subsection{Information Set Generation}
The concept of information sets is perhaps best explained through an example. 
Figure~\ref{listing} and show a full listing and the associated pictorial progression of an actual
kriegspiel game~\cite{li94chess}.  We have adapted the notation used by Wolfe for Berkeley Kriegspiel.  Odd (resp. even) moves are for white (resp. black).  The actual move is shown
first followed by a list of referee announcements.  Captures announcements are prefixed with an X and give the location
of the captured piece.  Similarly, pawn tries announcements are prefixed with T.  Attempted moves that were ultimately
declared illegal are shown in a list following the `:'.  For example at move 9, (Figure~\ref{fig:9W}--\ref{fig:9B}),
black attempts to block a potential threat along the diagonal by moving a pawn to f6, then attempts to escape or capture
a threat at f7.  When both of these attempts fail, black can infer that his king is being threatened by a protected
bishop or queen at f7.  The attempt to move to e7 is black's way of finding out whether the threatening piece is a queen
or bishop.  When this move fails, black knows it is white's queen and ultimately retreats to d8.  Note that the failed
moves by white on turns 3 and 4 are a result of being blocked by the pawn at e2.

@@

{The Information Set Generation Algorithm}
ISG can be formulated as a combinatorial search problem.  The input to the algorithm is a list
of observations for one player in a kriegspiel game.

@@

\begin{figure}[htpb]
\small
%\begin{algorithm}
\begin{boxedminipage}{\columnwidth}
\begin{algorithmic}[1]
\Function{DFS-Infoset}{$a_{1:t}$, $o_{1:T}^{(k)}$, $I$, $s$, $t$}
  \If {$t > T$}
    \State {$I \leftarrow I \cup a_{1:t}$} %\Comment{$a_{1:t}$ leads to $o_{1:T}$ for $P_k$; add it to the information set.}
  \Else {
    \For{\textbf{each} legal action $a$ at $s$}
      \State {$s'$ = {\sc GetNextState}($s$,$a$)}\label{line:update}
      \If {$s'.o_{t+1}^{(k)}=o_{t+1}^{(k)}$}
        \State {{\sc DFS-Infoset}($a_{1:t}a$, $o_{1:T}^{(k)}$, $I$, $s'$, $t+1$)} 
          %\Comment{ Recall that $N(a_{1:t}a) = T(T(a_{1:t}),a)$}
      \EndIf
    \EndFor
  }\EndIf
\EndFunction
\label{codelisting}
\end{algorithmic}
\end{boxedminipage}
\end{figure}

\Section{Charm++ and ParSSSE}\label{ParSSSE}
We built our search engine on top of the {\sc Charm++} run-time
system for parallel processing~\cite{kale93charm,kale09charm}.  {\sc Charm++}
is an extension to the C++ programming language that provides
machine-independent infrastructure for parallel computation.  The language and
its accompanying runtime system have been ported to many shared-memory and
distributed-memory platforms.

@@

\Section{Conclusion Future Work}
We have shown that information set generation for kriegspiel can be efficiently parallelized in a way that scales up to at least 1024 processors.

A key line of future research is to use our ISG routine as part of a full Kriegspiel play.  The ISG method itself could be updated using constraint-satisfaction techniques outlined in ~\cite{richards12information}, which improve pruning performance in the search space by incorporating information from all observations in the list at each search step, instead of performing depth-first search.   The use of ISG-based opponent modeling to improve overall decision making is described in ~\cite{richards12reasoning}.

@@

\Section{Conclusion}
We have shown that information set generation for kriegspiel can be efficiently parallelized.  The opportunities for
parallelism increase as the problem size increases.  Our algorithm has performed well on problems with information sets
with as many as 65 million nodes---much larger than any other treatment of the problem that we are aware of.  Because
the algorithm can be efficiently parallelized, it would be reasonable to explore using it in a full kriegspiel player.
It remains to be seen whether these scalability properties hold over a wide variety of playing styles and problem
instances.  We suspect that fruitful research remains with respect to alternative algorithms that implement variable
ordering heuristics.

>>>>

\author{Mark Richards, Abhishek Gupta, Osman Sarood and Laxmikant Kale\\
Department of Computer Science\\ 
University of Illinois\\ 
Urbana, IL\\ {mdrichar, gupta59, sarood1, kale}@illinois.edu\\
% For a paper whose authors are all at the same institution, 
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'', 
% just like the second author.
%\and
%Second Author\\
%Institution2\\
%First line of institution2 address\\ Second line of institution2 address\\ 
%SecondAuthor@institution2.com\\
}

@@

%\Section{Information Set Generation}
\Section{Background: Kriegspiel}
Information Set Generation is a widely adopted operation used in game trees search. Here, we present our techniques and results in the context of a particular game - Kriegspiel. However, the techniques and framework presented in this paper is general and can be applied to other imperfect games for generating information set.
%\Section{Background: Kriegspiel}
Kriegspiel, or partially observable chess,  was invented by Henry Michael Temple in 1899.  Historically, the
game has required three parties: two competing players and one referee.  Today,
players often compete over a computer network, with the role of the referee
being automated.  One player sees and controls only the white pieces; the
other, black.  The referee track of both sets of pieces.  The players alternate
moves as in a regular chess game, but instead of openly declaring their moves,
they secretly communicate move requests to the referee.  The referee, who knows
the location of all pieces, checks the legality of the move.

@@

%\SubSection{Example}
\begin{figure}[ht]
\centering
\begin{minipage}[b]{0.5\linewidth}
\includegraphics[width=0.7\textwidth]{images/9B.png}
\caption{Actual Position}
\label{fig:figure1}
\end{minipage}
\hspace{0.5cm}
\centering
\begin{minipage}[b]{0.5\linewidth}
\includegraphics[width=0.7\textwidth]{images/OneView.png}
\caption{White's view}
\label{fig:figure2}
\end{minipage}
\end{figure}

@@

\Section{Information Set Generation}
%\subsection{Information Set Generation}
The concept of information sets is perhaps best explained through an example. 
Figure~\ref{listing} and show a full listing and the associated pictorial progression of an actual
kriegspiel game~\cite{li94chess}.  We have adapted the notation used by Wolfe for Berkeley Kriegspiel.  Odd (resp. even) moves are for white (resp. black).  The actual move is shown
first followed by a list of referee announcements.  Captures announcements are prefixed with an X and give the location
of the captured piece.  Similarly, pawn tries announcements are prefixed with T.  Attempted moves that were ultimately
declared illegal are shown in a list following the `:'.  For example at move 9, (Figure~\ref{fig:9W}--\ref{fig:9B}),
black attempts to block a potential threat along the diagonal by moving a pawn to f6, then attempts to escape or capture
a threat at f7.  When both of these attempts fail, black can infer that his king is being threatened by a protected
bishop or queen at f7.  The attempt to move to e7 is black's way of finding out whether the threatening piece is a queen
or bishop.  When this move fails, black knows it is white's queen and ultimately retreats to d8.  Note that the failed
moves by white on turns 3 and 4 are a result of being blocked by the pawn at e2.

@@

\textbf{The Information Set Generation Algorithm}
(ISG) can be formulated as a combinatorial search problem.  The input to the algorithm is a list
of observations for one player in a kriegspiel game.

@@

\begin{figure}[htpb]
\small
%\begin{algorithm}
\begin{boxedminipage}{\columnwidth}
\begin{algorithmic}[1]
\Function{DFS-Infoset}{$a_{1:t}$, $o_{1:T}^{(k)}$, $I$, $s$, $t$}
  \If {$t > T$}
    \State {$I \leftarrow I \cup a_{1:t}$} %\Comment{$a_{1:t}$ leads to $o_{1:T}$ for $P_k$; add it to the information set.}
  \Else {
    \For{\textbf{each} legal action $a$ at $s$}
      \State {$s'$ = {\sc GetNextState}($s$,$a$)}\label{line:update}
      \If {$s'.o_{t+1}^{(k)}=o_{t+1}^{(k)}$}
        \State {{\sc DFS-Infoset}($a_{1:t}a$, $o_{1:T}^{(k)}$, $I$, $s'$, $t+1$)} 
          %\Comment{ Recall that $N(a_{1:t}a) = T(T(a_{1:t}),a)$}
      \EndIf
    \EndFor
  }\EndIf
\EndFunction
\label{codelisting}
\end{algorithmic}
\end{boxedminipage}
\caption{Algorithm}
\end{figure}

\Section{Parallel Implementation using Charm++ and ParSSSE}\label{ParSSSE}
We built our search engine on top of the {\sc Charm++} run-time
system for parallel processing~\cite{kale93charm,kale09charm}.  {\sc Charm++}
is an extension to the C++ programming language that provides
machine-independent infrastructure for parallel computation.  The language and
its accompanying runtime system have been ported to many shared-memory and
distributed-memory platforms.

@@

%\Section{Conclusion Future Work}
%We have shown that information set generation for kriegspiel can be efficiently parallelized in a way that scales up to at least 1024 processors.

%A key line of future research is to use our ISG routine as part of a full Kriegspiel play.  The ISG method itself could be updated using constraint-satisfaction techniques outlined in ~\cite{richards12information}, which improve pruning performance in the search space by incorporating information from all observations in the list at each search step, instead of performing depth-first search.   The use of ISG-based opponent modeling to improve overall decision making is described in ~\cite{richards12reasoning}.

@@

\Section{Conclusion and Future Work}
We have shown that information set generation for kriegspiel can be efficiently parallelized.  The opportunities for
parallelism increase as the problem size increases.  Our algorithm has performed well on problems with information sets
with as many as 65 million nodes---much larger than any other treatment of the problem that we are aware of.  Because
the algorithm can be efficiently parallelized, it would be reasonable to explore using it in a full kriegspiel player.
It remains to be seen whether these scalability properties hold over a wide variety of playing styles and problem
instances.  We suspect that fruitful research remains with respect to alternative algorithms that implement variable
ordering heuristics.
